\chapter{Extension to Maximum Matching}
\label{chap:maximum_matching}

This chapter shows how to extend the perfect matching algorithms to solve the maximum matching problem.

\section{Maximum Matching algorithm}

The extension to maximum matching is based on the following theorem by Lovasz.
\begin{theorem}[\citet{plummer1986matching}]
\label{thm:rank_matching}
    Let \(G\) be a graph and let \(T\) be the Tutte Matrix of \(G\).
    Then, \(\rank(T_G) = 2\nu(G)\).
\end{theorem}

\begin{proof}
TODO - Prove.
\end{proof}

According to \cref{thm:rank_matching}, the number of unmatched vertices can be directly computed as:
\[
    |V(G)| - \rank(T).
\]
To address these unmatched vertices, we can construct an augmented graph by introducing new vertices strategically connected to all existing vertices. 
Specifically, for each unmatched vertex in the original graph, we add a new vertex with connections to every vertex in the original graph. 
This transformation ensures that every previously unmatched vertex now has an adjacent vertex, thus ensuring the existence of a perfect matching.

\begin{programruledcaption}{Maximum Matching algorithm}
    \begin{lstlisting}[
      language={pseudocode},
      style=pseudocode,
      style=wider,
      functions={PerfectMatching, TutteMatrix, Rank},
      specialidentifiers={},
    ]
        function MaximumMatching(G)
            $T$ := TutteMatrix(G) // Tutte matrix of \(G\) with random values.
            $G'$ := $G$
            for i := 0; i < |V(G)| - rank(G); i := i + 1 
                $v$ := new vertex
                $V(G')$ := $V(G') \cup \{v\}$
                for $u \in V$ do 
                    $E(G')$ := $E(G') \cup \{uv\}$
                end
            end
            $M'$ := PerfectMatching($G'$)
            $M$ := $\emptyset$
            for $uv \in M'$ do 
                if $u \in E$ and $v \in V$ then // If this edge exists in the original graph.
                    $M$ := $M \cup \{uv\}$
                end
            end
            return M
        end
    \end{lstlisting}
\end{programruledcaption}
\noindent
\textbf{Time Complexity Analysis:} Let \(t(G)\) be the total running time of \(\SC{MaximumMatching}(G)\) and \(f(G')\) be the running time of \(\SC{MaximumMatching}(G')\) on the modified graph.
\begin{itemize}
\item \textbf{Augmented Graph creation (Lines 4 to 10)}:
The algorithm identifies unmatched vertices and adds corresponding new vertices. 
Since there are at most \(n\) unmatched vertices and each new vertex connects to \(n\) vertices, this phase requires \(O(n^2)\) time;

\item \textbf{Perfect Matching in the augmented graph (Line 11)}: 
The augmented graph has at most \(2n\) vertices. 
By \cref{alg:harvey_complexity}, this step has a time complexity  \(f(G') = O((2n)^\omega) = O(2^\omega n^\omega) = O(n^\omega)\);

\item \textbf{Maximum Matching recovery (Lines 13 to 17)}: 
Checking whether a vertex belongs to the original graph can be implemented in \(O(1)\) time by using integer indexing. 
The overall complexity for this verification across all vertices is \(O(n)\).
\end{itemize}
Combining these steps, the total time complexity is:
\[
t(G) = O(n^2) + f(G') + O(n^2) = O(n^2) + O(n^\omega) + O(n^2) = O(n^\omega). \numberthis \label{alg:maximum_matching_complexity}
\]

\section{Analysis}